---
title: "Programming in R Assignment"
author: "Sherine"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
#Install the ggextra library
install.packages("ggExtra",repos = "http://cran.us.r-project.org")
#Load the tidyverse and ggextra libraries
library(tidyverse)
library("ggExtra")
```

## Including Plots

Recreate the scatterplot of the lab 1 exercise of class 1:

```{r scatterplot, message=FALSE}
ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy, color = as.factor(cyl))) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color = "brown",
              se = FALSE,
              linewidth = 0.5) +
  labs(
    x = "City MPG",
    y = "Highway MPG",
    title = "Highway MPG vs. City MPG",
    color = "N. of cylinders"
  )

```

Move the position of the legend to the bottom of the plot

```{r scatterplot2, message=FALSE}
ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy, color = as.factor(cyl))) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color = "brown",
              se = FALSE,
              linewidth = 0.5) +
  labs(
    x = "City MPG",
    y = "Highway MPG",
    title = "Highway MPG vs. City MPG",
    color = "N. of cylinders"
  )+ theme(legend.position="bottom")
```

Use the ggMarginal function from the ggExtra library to add marginal histograms for both the x and y axis of your scatterplot


```{r scatterplot3, message=FALSE}
p <- ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy, color = as.factor(cyl))) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color = "brown",
              se = FALSE,
              linewidth = 0.5) +
  labs(
    x = "City MPG",
    y = "Highway MPG",
    title = "Highway MPG vs. City MPG",
    color = "N. of cylinders"
  )+ theme(legend.position="bottom")
ggMarginal(p,type="histogram")
```

Configure both marginal histograms to have 16 bins and to be lightblue in color
```{r scatterplot4, message=FALSE}
p <- ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy, color = as.factor(cyl))) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color = "brown",
              se = FALSE,
              linewidth = 0.5) +
  labs(
    x = "City MPG",
    y = "Highway MPG",
    title = "Highway MPG vs. City MPG",
    color = "N. of cylinders"
  )+ theme(legend.position="bottom")
ggMarginal(p,type="histogram",bins=16,fill="lightblue")
```

Make the size of the plot 10in x 8in

Added `fig.dim=c(10,8)` to change the size of the plot making width 10 and height 8
```{r scatterplot5, message=FALSE,fig.dim=c(10,8)}

p <- ggplot(data = mpg,
       mapping = aes(x = cty, y = hwy, color = as.factor(cyl))) +
  geom_jitter(alpha = 0.4) +
  geom_smooth(color = "brown",
              se = FALSE,
              linewidth = 0.5) +
  labs(
    x = "City MPG",
    y = "Highway MPG",
    title = "Highway MPG vs. City MPG",
    color = "N. of cylinders"
  )+ theme(legend.position="bottom")
ggMarginal(p,type="histogram",bins=16,fill="lightblue")

```

